Site location study for a high-mountain millimeter observatory in the Colombian Andes I: In situ data

### 2017-04-20 pasp/paperclimate.tex
more plots in paper

Besides astronomical observations, satellite communications technology in the THz frequency range is a rapidly growing field \cite{newera}. Atmospheric models for absorption of THz photons \cite{rosenkranz,lababs} and artifact removal models \cite{removal} have been developed recently, specially considering that the ``classical radio window'' in telecommunications has become saturated, and terabit satellite links could be used in the near future \cite{suen2016}. Near-equatorial countries face a challenge in using this frequency band for satellite telecommunications and astronomical observations due to the presence of a tropical belt of dense water vapor, which efficiently absorbs THz radiation \cite{tropicalbelt}. Previous studies of astronomical site testing in the visible range have yielded that nighttime humidity conditions across the northern Andes are only suitable for educational observatories \cite{pinzon}. However, considering that millimeter/sub-millimeter observations need not be done during the night, unusually dry, clear-daytime-sky, high altitude regions in the Northern Andes could be suitable for a world-class millimeter-wave observatory. \\

>>>>

Besides astronomical observations in the THz frequency range, development of THz satellite communications technology is also a rapidly growing field \cite{newera}. Atmospheric models for absorption of THz photons \cite{rosenkranz,lababs} and artifact removal models \cite{removal} have been developed recently, specially considering that the ``classical radio window'' in telecommunications has become saturated, and terabit satellite links could be used in the near future \cite{suen2016}. Near-equatorial countries face a challenge in using this frequency band for satellite telecommunications and astronomical observations due to the presence of a tropical belt of dense water vapor, which efficiently absorbs THz radiation \cite{tropicalbelt}. Previous studies of astronomical site testing in the visible range have yielded that nighttime humidity conditions across the northern Andes are only suitable for educational observatories \cite{pinzon}. However, considering that millimeter/sub-millimeter observations need not be done during the night, unusually dry, clear-daytime-sky, high altitude regions in the Northern Andes could be suitable for a world-class millimeter-wave observatory. \\
